*Prochlorococcus marinus* responses to light and oxygen

### 2024-03-07 Code/Manuscript_O2.Rmd
minor edits

```{r load libraries, include = FALSE}
# libraries; Note check actual dependencies
# Naaman add packages?
library(knitr)
library(tidyverse)
library(broom)
library(grid)
library(bookdown)
library(glue)
library(cowplot)
library(ggpubr)

library(png)
library(cowplot)

@@

*Prochlorococcus marinus* comprises many strains, organized into clades, defined by 16S-23S intergenic transcribed ribosomal sequence signatures [@rocapGenomeDivergenceTwo2003]. The clades inhabit distinct ecological niches [@moorePhysiologyMolecularPhylogeny1998], originally defined as High-Light (HL) or Low-Light (LL).  Of at least 12 known *Prochlorococcus* genetic clades only 5 to date have cultured representatives; HLI, HLII, LLI, LLII/III and LIV [@billerProchlorococcusStructureFunction2015]. Current niches of *P. marinus* strains span ocean water columns [@veldhuisVerticalDistributionPigment1990; @chisholmNovelFreelivingProchlorophyte1988; @partenskyProchlorococcusMarinePhotosynthetic1999] and extend into regions with low dissolved oxygen concentrations [@partenskyProchlorococcusAdvantagesLimits2010; @goerickeNovelNicheProchlorococcus2000; @lavinNovelLineagesProchlorococcus2010; @ulloaCyanobacteriumProchlorococcusHas2021; @garcia-robledoCrypticOxygenCycling2017].

@@

HL clades are more recently branching lineages, with reduced genome sizes in comparison to LL clades. High-Light clades are typically dominant picophytoplankters in near-surface, oligotrophic waters, characterized by high light levels. Clade HLI, represented by cultured strain MED4, is adapted to higher iron, and lower temperatures, and originated from 5 m depth in the Mediterranean Sea [@partenskyProchlorococcusMarinePhotosynthetic1999]. Clade HLII, adapted to higher iron, and higher temperatures, is the most abundant *P. marinus* clade in the North Atlantic and North Pacific Oceans, often constituting over 90% of the total population [@partenskyProchlorococcusMarinePhotosynthetic1999], and are most numerous around 50 m depth [@partenskyProchlorococcusMarinePhotosynthetic1999]. Clade HLIII/IV is adapted to lower iron [@kentParallelPhylogeographyProchlorococcus2019; @johnsonNichePartitioningProchlorococcus2006; @zinserInfluenceLightTemperature2007].

@@

We used the OceansMap Protein Portal (OPP; https://proteinportal.whoi.edu/) [@saitoDevelopmentOceanProtein2021] to analyze the distribution of proteins from clades of *P. marinus* in samples taken across a range of [O~2~] and depth, which in turn correlates to peak light at the site of sampling. In parallel we analyzed the growth and physiological responses of representative strains from three clades of *P. marinus* under a matrix of [O~2~], light levels, spectral waveband ranges, and photoperiods, to approximate eco-physiological conditions representative of current and hypothetical future ocean zones. *Prochlorococcus marinus* MED4, a clade HLI strain, was isolated near the ocean surface (5m depth) of the Mediterranean Sea where [O~2~] is near saturation, light levels are high and spectral bias from full solar irradiance is minimal. *Prochlorococcus marinus* SS120, a clade LLII/III strain, was isolated from the Sargasso Sea at a depth of 120 m, while *P. marinus* MIT9313, a clade LLIV strain, was isolated from the North Atlantic Gulf Stream at a depth of 135 m. At these depths, light attenuation and spectral shifts occur, resulting in low blue light, while [O~2~] varies from near-surface saturation levels to decreased concentrations, but does not necessarily decrease systematically with depth [@billheimerOxygenSeasonalityUtilization2021].

@@

PAR of 180, 90 or 30 µmol photons m^-2^ s^-1^, and spectral wavebands (white LED full spectrum, 660 nm, and 450 nm) were chosen to approximate light
levels and spectral colours spanning the vertical ocean water column, from near-surface to the lower euphotic zone depths. Photoperiods were chosen to approximate diel cycles characteristic of current and hypothetical future niches of *P. marinus*; 16 h represents temperate (45°N) summer at the ocean surface; 12 h for equatorial (0°N)
ocean surface or temperate (45°N) spring and fall ocean surface or temperate (45°N) summer at deeper ocean depths; 8 h for temperate (45°N) winter at the surface or at temperate (45°N) spring and fall at depth and equatorial (0°N) deep ocean depths; and 4 h for temperate (45°N) winter or deep ocean depths during temperate (45°N) spring and fall.

>>>>

```{r load libraries, include = FALSE}
# libraries; Note check actual dependencies
# Naaman add packages?
library(knitr)
library(tidyverse)
library(broom)
library(grid)
library(bookdown)
library(glue)
library(cowplot)
library(ggpubr)
library(png)
library(cowplot)

@@

*Prochlorococcus marinus* comprises many strains, organized into clades, defined by 16S-23S intergenic transcribed ribosomal sequence signatures [@rocapGenomeDivergenceTwo2003]. The clades inhabit distinct ecological niches [@moorePhysiologyMolecularPhylogeny1998], originally defined as High-Light (HL) or Low-Light (LL). Only 5 out of 12 known *Prochlorococcus* genetic clades have cultured representatives to date; HLI, HLII, LLI, LLII/III and LIV [@billerProchlorococcusStructureFunction2015]. Current niches of *P. marinus* strains span ocean water columns [@veldhuisVerticalDistributionPigment1990; @chisholmNovelFreelivingProchlorophyte1988; @partenskyProchlorococcusMarinePhotosynthetic1999] and extend into regions with low dissolved oxygen concentrations [@partenskyProchlorococcusAdvantagesLimits2010; @goerickeNovelNicheProchlorococcus2000; @lavinNovelLineagesProchlorococcus2010; @ulloaCyanobacteriumProchlorococcusHas2021; @garcia-robledoCrypticOxygenCycling2017].

@@

HL clades are more recently branching lineages, with reduced genome sizes in comparison to LL clades. High-Light clades are typically dominant picophytoplankters in near-surface, oligotrophic waters, characterized by high light levels and subdivided into clades based on iron adaptation [@johnsonNichePartitioningProchlorococcus2006; @zinserInfluenceLightTemperature2007; @kentParallelPhylogeographyProchlorococcus2019]. Clade HLI, represented by cultured strain MED4, is adapted to high iron, and low temperatures, and originated from 5 m depth in the Mediterranean Sea [@partenskyProchlorococcusMarinePhotosynthetic1999]. Clade HLII, adapted to high iron, and high temperatures, is the most abundant *P. marinus* clade in the North Atlantic and North Pacific Oceans, often constituting over 90% of the total population [@partenskyProchlorococcusMarinePhotosynthetic1999], and are most numerous around 50 m depth [@partenskyProchlorococcusMarinePhotosynthetic1999]. Clade HLIII/IV is adapted to low iron [@kentParallelPhylogeographyProchlorococcus2019; @johnsonNichePartitioningProchlorococcus2006; @zinserInfluenceLightTemperature2007].

@@

We used the OceansMap Protein Portal (OPP; https://proteinportal.whoi.edu/) [@saitoDevelopmentOceanProtein2021] to analyze the distribution of proteins from clades of *P. marinus* in samples taken across a range of [O~2~] and depth, which in turn correlates to peak light at the site of sampling. In parallel we analyzed the growth and physiological responses of representative strains from three clades of *P. marinus* under a matrix of [O~2~], light levels, spectral waveband ranges, and photoperiods, to approximate eco-physiological conditions representative of current and hypothetical future ocean zones. *Prochlorococcus marinus* MED4, a clade HLI strain, was isolated near the ocean surface (5m depth) of the Mediterranean Sea where [O~2~] is near saturation, light levels are high and spectral bias from full solar irradiance is minimal. *Prochlorococcus marinus* SS120, a clade LLII/III strain, was isolated from the Sargasso Sea at a depth of 120 m, while *P. marinus* MIT9313, a clade LLIV strain, was isolated from the North Atlantic Gulf Stream at a depth of 135 m [@moorePhotophysiologyMarineCyanobacterium1999]. At these depths, light attenuation and spectral shifts occur, resulting in low blue light, while [O~2~] varies from near-surface saturation levels to decreased concentrations, but does not necessarily decrease systematically with depth [@billheimerOxygenSeasonalityUtilization2021].

@@

PAR of 180, 90 or 30 µmol photons m^-2^ s^-1^, and spectral wavebands (white LED full spectrum, 660 nm (red light), and 450 nm (blue light)) were chosen to approximate light
levels and spectral colours spanning the vertical ocean water column, from near-surface to the lower euphotic zone depths. Photoperiods were chosen to approximate diel cycles characteristic of current and hypothetical future niches of *P. marinus*; 16 h represents temperate (45°N) summer at the ocean surface; 12 h for equatorial (0°N)
ocean surface or temperate (45°N) spring and fall ocean surface or temperate (45°N) summer at deeper ocean depths; 8 h for temperate (45°N) winter at the surface or at temperate (45°N) spring and fall at depth and equatorial (0°N) deep ocean depths; and 4 h for temperate (45°N) winter or deep ocean depths during temperate (45°N) spring and fall.
